---
title: "Imaging, shape, and volumetric workflows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Imaging, shape, and volumetric workflows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6.5, fig.height = 4.5)
library(SpatialMethodsWorkbench)
```

This vignette uses three compact deterministic datasets: 48 image subjects, six contours, and 72 volumetric observations.

## Wide-image scalar regression screen

```{r image-data}
image <- demo_image_data(n = 48, features = 8, seed = 2026)
pixel_features <- grep("^pixel_", names(image), value = TRUE)
```

```{r scalar-image}
scalar <- run_analysis(
  "scalar_image", image, list(outcome = "outcome"), pixel_features,
  list(image_ridge = 5, seed = 2026)
)
scalar$summary
head(scalar$table)
```

```{r scalar-image-plot}
scalar$plot
```

## Latent image-to-image regression screen

```{r image-to-image}
paired_features <- grep("^(input__|output__)", names(image), value = TRUE)
latent <- run_analysis(
  "image_to_image", image, list(), paired_features,
  list(latent_factors = 3L, seed = 2026)
)
latent$summary
head(latent$table)
```

```{r image-to-image-plot}
latent$plot
```

These fast screens expect already aligned and consistently ordered image features. They are not the full SV-NN, ST-CAR, or SBLF posterior engines.

```{r image-method-citations}
method_citations("scalar_image")
method_citations("image_to_image")
```

## Ordered-contour shape PCA

```{r contour-data}
contours <- demo_contour_data(contours = 6, points = 16, seed = 2026)
head(contours)
```

```{r shape-pca}
shape <- run_analysis(
  "shape_pca", contours,
  list(id = "contour_id", x = "x", y = "y"), character(),
  list(landmarks = 16L)
)
shape$summary
shape$table
```

```{r shape-pca-plot}
shape$plot
```

Input points must follow each boundary in order. The vignette reports reproducible scores without assigning a tumor phenotype.

## Volumetric ISPAT-3D input

```{r volume-data}
volume <- demo_spatial_3d_data(n = 72, volumes = 2, seed = 2026)
volume_features <- c("epithelial", "APC", "CTL", "Treg", "T_helper")
volume_mapping <- list(
  x = "x", y = "y", z = "z", zone = "zone", id = "volume_id"
)
validate_dataset(volume, volume_mapping)[c("ok", "n", "p")]
```

The public ISPAT-3D engine is opt-in and is not executed during vignette builds.

```{r ispat3d, eval=FALSE}
remotes::install_github("sagnikbhadury/ISPAT-3D")

ispat3d <- run_analysis(
  "ispat3d", volume, volume_mapping, volume_features,
  list(kernel = "Matern", factor_fit = "CAVI", threshold = 0.15, cores = 1L)
)
ispat3d$summary
head(ispat3d$table)
```

The 72-row dataset is a software demonstration. Registered coordinates, section alignment, zones, and study-specific sampling must be established before using real volumetric data.
