---
title: "Prediction, clustering, and mediation workflows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Prediction, clustering, and mediation workflows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6.5, fig.height = 4.5)
library(SpatialMethodsWorkbench)
```

All examples use the same 72-row synthetic table and fixed seed. Small settings make the vignette suitable for a laptop or continuous integration.

These executable workflows are adaptations informed by the public AI for Beginners, ML for Beginners, and Deep Understanding of Deep Learning repositories credited by `method_citations()`. Those curriculum repositories are not imported as R packages and are not presented as statistical engines.

```{r data}
d <- demo_spatial_data(n = 72, seed = 2026)
features <- c("epithelial", "APC", "CTL", "Treg", "T_helper")
mapping <- list(
  x = "x", y = "y", z = NULL, zone = "zone", id = "sample_id",
  outcome = "disease", exposure = "epithelial", mediator = "APC"
)
```

## Held-out spatial prediction

```{r spatial-ml}
prediction <- run_analysis(
  "spatial_ml", d, mapping, features,
  list(seed = 2026)
)
prediction$summary
head(prediction$table)
```

```{r spatial-ml-plot}
prediction$plot
```

## Spatially weighted clustering

```{r clustering}
clusters <- run_analysis(
  "spatial_clustering", d, mapping, features,
  list(clusters = 3L, spatial_weight = 0.5, seed = 2026)
)
clusters$summary
head(clusters$table)
```

```{r clustering-plot}
clusters$plot
```

## Shallow-neural prediction

The `nnet` package is a suggested dependency. The model below has only four hidden units and a small dataset.

```{r neural}
if (requireNamespace("nnet", quietly = TRUE)) {
  neural <- run_analysis(
    "neural_prediction", d, mapping, features,
    list(hidden_units = 4L, decay = 0.02, seed = 2026)
  )
  neural$summary
}
```

```{r neural-plot}
if (exists("neural")) neural$plot
```

## Mediation with a small bootstrap

For this synthetic execution check, `epithelial` is mapped to exposure, `APC` to mediator, and `CTL` to outcome. Twenty-five bootstrap replicates keep rendering fast; a real analysis requires settings justified for the study.

```{r mediation}
mediation_mapping <- mapping
mediation_mapping$outcome <- "CTL"
mediation <- run_analysis(
  "mediation", d, mediation_mapping, features,
  list(bootstrap = 25L, seed = 2026)
)
mediation$table
```

```{r mediation-plot}
mediation$plot
```

These workflows return predictive, clustering, and model-based mediation quantities. The code demonstrates execution and object structure; it does not validate biological labels, clinical utility, or causal identification.

```{r fork-citations}
method_citations("spatial_ml")
method_citations("neural_prediction")
```
