---
title: "Auditing reproducibility bundles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Auditing reproducibility bundles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6.5, fig.height = 4.5)
library(SpatialMethodsWorkbench)
```

Every successful app or programmatic run can be exported as a reproducibility bundle. This example uses 60 synthetic observations, 19 permutations, and a fixed seed.

```{r run}
d <- demo_spatial_data(n = 60, seed = 2026)
mapping <- list(x = "x", y = "y")
features <- "APC"
result <- run_analysis(
  "spatial_qc", d, mapping, features,
  list(feature = "APC", neighbors = 5L, permutations = 19L, seed = 2026)
)
```

## Inspect the manifest before writing

```{r manifest}
manifest <- result_manifest(result)
manifest[c("application", "application_version", "analysis_id", "method", "features")]
manifest$parameters
```

## Write and list the archive

```{r write}
bundle <- file.path(tempdir(), "spatial-methods-workbench-bundle.zip")
write_result_bundle(result, bundle)
contents <- utils::unzip(bundle, list = TRUE)
contents[, c("Name", "Length")]
```

## Audit the archived metadata

```{r audit}
extract_dir <- tempfile("workbench-bundle-")
dir.create(extract_dir)
utils::unzip(bundle, exdir = extract_dir)

archived_manifest <- jsonlite::read_json(
  file.path(extract_dir, "manifest.json"), simplifyVector = TRUE
)
archived_manifest[c("application_version", "analysis_id", "method", "features")]

readLines(file.path(extract_dir, "CITATION.txt"), n = 8)
```

The archive should contain `results.csv`, `figure.png`, `manifest.json`, `result_object.rds`, `CITATION.txt`, and `COLLABORATION.txt`. It intentionally excludes the input dataset and automated result interpretation.

## Reopen the exact R result

```{r reopen}
restored <- readRDS(file.path(extract_dir, "result_object.rds"))
identical(restored$analysis_id, result$analysis_id)
identical(restored$params, result$params)
```

Record the Workbench version, method citations, seed, mapping, features, and controls when reporting an analysis. Scientific interpretation still requires study design and domain context outside the application.
