---
title: "Spatial diagnostics and network workflows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Spatial diagnostics and network workflows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6.5, fig.height = 4.5)
library(SpatialMethodsWorkbench)
```

This vignette exercises the spatial and network analysis family with 72 synthetic observations. The small permutation count and compact data are intended for reproducibility checks, not final analysis settings.

```{r data}
d <- demo_spatial_data(n = 72, seed = 2026)
features <- c("epithelial", "APC", "CTL", "Treg", "T_helper")
mapping <- list(
  x = "x", y = "y", z = NULL, zone = "zone", id = "sample_id",
  outcome = "disease", exposure = "epithelial", mediator = "APC"
)
```

## Spatial exploration and autocorrelation

```{r spatial-qc}
spatial_qc <- run_analysis(
  "spatial_qc", d, mapping, features,
  list(feature = "epithelial", neighbors = 6, permutations = 19, seed = 2026)
)
spatial_qc$table
```

```{r spatial-qc-plot}
spatial_qc$plot
```

The output reports the configured neighbor graph, statistic, permutation calculation, and provenance. It does not establish a biological mechanism.

## Region-aware conditional network

```{r conditional-network}
conditional <- run_analysis(
  "conditional_network", d, mapping, features,
  list(threshold = 0.15, ridge = 0.05, by_zone = TRUE)
)
head(conditional$table, 10)
conditional$summary
```

```{r conditional-network-plot}
conditional$plot
```

The stored matrices can be audited directly.

```{r matrix-names}
names(conditional$matrices)
vapply(conditional$matrices, nrow, integer(1))
```

Edges are regularized conditional-association summaries. The vignette intentionally stops at reproducible output inspection.

## Public advanced engines

The following calls use the same 72-row table. They are not evaluated while building the vignette because runtime depends on the installed public engines and local hardware.

### ISPat 2D

```{r ispat, eval=FALSE}
remotes::install_github("sagnikbhadury/ISPAT")

ispat <- run_analysis(
  "ispat", d, mapping, features,
  list(
    kernel = "Matern", spatial_fit = "VB", factor_fit = "CAVI",
    cores = 1L, threshold = 0.15
  )
)
ispat$summary
head(ispat$table)
```

### GP-GHS

The short chain below is a software smoke check. Increase Monte Carlo settings and perform convergence assessment for research use.

```{r gpghs, eval=FALSE}
remotes::install_github("sagnikbhadury/GP-GHS")

gpghs <- run_analysis(
  "gpghs", d, mapping, features,
  list(
    basis = 3L, nu = 1.5, nmc = 200L, burn = 50L, thin = 2L,
    threshold = 0.15, symmetry = "AND", cores = 1L
  )
)
gpghs$summary
head(gpghs$table)
```

Use `method_citations("ispat")` and `method_citations("gpghs")` to retrieve the required software and method references.
